---
title: "Protein"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Protein}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In our analysis, we'll gather structural data from PolBase and KLIFS databases, focusing on DNA polymerases and kinases, respectively. We'll classify proteins into families and compare their structures, plotting amino acid proportions and analyzing sequence lengths. Our R package will streamline data cleaning, leveraging Tidyverse functions and a protein class to enhance organization and reusability.

Challenges may arise in parsing complex PDB files, but we'll address them using filtering techniques and substring searches. Our analysis offers educational value, providing people with insights into protein structure-function relationships and empowering them to explore and compare protein structures using our R package.

# Installation

Before installing the `protein5555` package, it's important to set a CRAN mirror. You can do this by executing the following code:

```{r}
# Set CRAN mirror
options(repos = "https://cloud.r-project.org")
```

To use the `protein5555` package, you'll need to install it from GitHub using the `devtools` package. Follow these steps:

```{r}
# Install devtools and bio3d package (if not already installed)
install.packages("devtools")
install.packages("bio3d")

# Load devtools and bio3d library
library(devtools)
library(bio3d)

# Install the protein5555 package from GitHub
devtools::install_github("EmelineHaroldsen/protein5555")
```

Once installed, you can load the protein5555 package in your R session using the following command:

```{r}
library(protein5555)
```

# Usage

In this section, we'll demonstrate how to use the `protein5555` package to analyze protein structures and perform comparative analyses.

# Example Data

# References

# Conclusion
